### 2020-11-11 paper/supplement.tex
First draft of paper

\documentclass{article}
\usepackage[lmargin=1 in, rmargin=1 in, tmargin=1 in, bmargin=1 in]{geometry}
\usepackage{lineno}
\linenumbers

\begin{document}

\title{Supplementary Material}
\date{}
\maketitle

\begin{table}
\caption{Temporary stations used for the 2007-2009 catalog}
\centering
\begin{tabular}{c c c c c c c c c}
\hline
Station & Network & Channels & Location & Data center & Latitude & Longitude & Begin time & End time \\
\hline
B039 & PB & EH1,EH2,EHZ & -- & IRIS & 41.4667 & -122.4847 & 2007-10-15 & 3000-01-01 \\
KCPB & NC & HHE,HHN,HHZ & -- & NCEDC & 39.6863 & -123.5824 & 2002-10-17 & 3000-01-01 \\
KHBB & NC & HHE,HHN,HHZ & -- & NCEDC & 40.6599 & -123.2197 & 2003-09-11 & 3000-01-01 \\
KRMB & NC & HHE,HHN,HHZ & -- & NCEDC & 41.5230 & -123.9080 & 2001-06-16 & 3000-01-01 \\
KSXB & NC & HHE,HHN,HHZ & -- & NCEDC & 41.8304 & -123.8769 & 2001-07-13 & 3000-01-01 \\
ME01 & XQ & BHE,BHN,BHZ & 01 & IRIS & 41.7752 & -123.4034 & 2007-07-31 & 2009-06-12 \\
ME02 & XQ & BHE,BHN,BHZ & 01 & IRIS & 41.6898 & -122.3372 & 2007-07-22 & 2009-05-20 \\
ME03 & XQ & BHE,BHN,BHZ & 01 & IRIS & 41.702 & -123.881 & 2007-09-27 & 2009-06-13 \\
ME04 & XQ & BHE,BHN,BHZ & 01 & IRIS & 41.355 & -123.514 & 2007-07-23 & 2009-06-12 \\
ME08 & XQ & BHE,BHN,BHZ & 01 & IRIS & 40.222 & -123.305 & 2007-07-22 & 2009-06-03 \\
ME09 & XQ & BHE,BHN,BHZ & 01 & IRIS & 41.086 & -122.726 & 2007-07-21 & 2009-05-14 \\
ME10 & XQ & BHE,BHN,BHZ & 01 & IRIS & 40.9591 & -122.4619 & 2007-10-31 & 2009-06-12 \\
ME12 & XQ & BHE,BHN,BHZ & 01 & IRIS & 40.104 & -122.498 & 2007-07-20 & 2009-06-12 \\
ME13 & XQ & BHE,BHN,BHZ & 01 & IRIS & 40.764 & -122.918 & 2007-07-22 & 2009-06-12 \\
ME16 & XQ & BHE,BHN,BHZ & 01 & IRIS & 40.577 &  -122.087 & 2007-07-20 & 2008-10-14 \\
ME24 & XQ & BHE,BHN,BHZ & 01 & IRIS & 40.359 & -122.047 & 2007-07-19 & 2009-05-15 \\
ME26 & XQ & BHE,BHN,BHZ & 01 & IRIS & 40.157 & -122.099 & 2007-07-26 & 2009-02-17 \\
ME27 & XQ & BHE,BHN,BHZ & 01 & IRIS & 40.453 & -123.155 & 2008-01-01 & 2009-06-13 \\
ME28 & XQ & BHE,BHN,BHZ & 01 & IRIS & 40.327 & -122.471 & 2007-07-20 & 2009-05-14 \\
ME29 & XQ & BHE,BHN,BHZ & 01 & IRIS & 41.142 & -123.1408 & 2007-09-21 & 2009-06-13 \\
ME30 & XQ & BHE,BHN,BHZ & 01 & IRIS & 40.843 & -123.565 & 2007-07-24 & 2009-06-22 \\
ME42 & XQ & BHE,BHN,BHZ & 01 & IRIS & 39.7145 & -123.273 & 2007-07-15 & 2009-06-16 \\
ME43 & XQ & BHE,BHN,BHZ & 01 & IRIS & 39.5703 & -123.175 & 2007-09-16 & 2009-06-16 \\
ME44 & XQ & BHE,BHN,BHZ & 01 & IRIS & 39.4928 & -123.187 & 2007-09-16 & 2009-06-15 \\
ME45 & XQ & BHE,BHN,BHZ & 01 & IRIS & 39.4666 & -122.9595 & 2007-09-12 & 2009-06-09 \\
ME46 & XQ & BHE,BHN,BHZ & 01 & IRIS & 39.196 & -122.968 & 2007-09-25 & 2009-06-11 \\
ME47 & XQ & BHE,BHN,BHZ & 01 & IRIS & 39.1124 & -122.8106 & 2007-09-25 & 2008-09-17 \\
ME49 & XQ & BHE,BHN,BHZ & 01 & IRIS & 39.8633 & -123.7194 & 2007-09-25 & 2009-06-08 \\
ME52 & XQ & BHE,BHN,BHZ & 01 & IRIS & 39.3228 & -123.2191 & 2008-02-21 & 2009-07-08 \\
ME54 & XQ & BHE,BHN,BHZ & 01 & IRIS & 39.0132 & -123.3779 & 2007-09-24 & 2009-03-03 \\
ME57 & XQ & BHE,BHN,BHZ & 01 & IRIS & 39.9118 & -122.5676 & 2007-10-24 & 2009-06-11 \\
WDC & BK & BHE,BHN,BHZ & -- & NCEDC & 40.5799 & -122.5411 &1992-09-17 & 2011-05-06 \\
YBH &  BK & BHE,BHN,BHZ & -- & NCEDC & 41.7320 & -122.7104 & 1993-07-24 & 2011-06-03 \\
\hline
\end{tabular}
\end{table}

\begin{table}
\caption{Permanent stations used for the 2004-2011 catalog}
\centering
\begin{tabular}{c c c c c c c c c}
\hline
Station & Network & Channels & Location & Data center & Latitude & Longitude & Begin time & End time \\
\hline
B039 & PB & EH1,EH2,EHZ & -- & IRIS & 41.4667 &  -122.4847 & 2007-10-15 & 3000-01-01 \\
B040 & PB & EH1,EH2,EHZ & -- & IRIS & 41.8308 &  -122.4205 & 2007-10-16 & 3000-01-01 \\
B933 & PB & EH1,EH2,EHZ & -- & IRIS & 40.0600 &  -123.9690 & 2008-09-13 & 3000-01-01 \\
B935 & PB & EH1,EH2,EHZ & -- & IRIS & 40.4787 &  -123.5732 & 2008-10-29 & 3000-01-01 \\
GASB & BK & BHE,BHN,BHZ & -- & NCEDC & 39.65471 & -122.71595 & 2005-09-22 & 2011-06-16 \\
GASB & BK & BHE,BHN,BHZ & 0 & NCEDC & 39.65471 & -122.71595 & 2011-06-16 & 3000-01-01 \\
GBB & NC & EHZ & -- & NCEDC & 39.80127 & -122.34550 & 2000-12-06 & 3000-01-01 \\
GHM & NC & EHZ & -- & NCEDC & 39.49545 & -122.93096 & 1984-01-01 & 3000-01-01 \\
GRO & NC & EHZ & -- & NCEDC & 39.91684 & -122.67117 & 1990-12-13 & 3000-01-01 \\
GSN & NC & EHZ & -- & NCEDC & 38.94040 & -123.19245 & 1984-01-01 & 2018-12-11 \\
GTC & NC & SHZ & -- & NCEDC & 39.39944 & -123.55532 & 1996-08-01 & 2011-10-27 \\
GVV & NC & EHZ & -- & NCEDC & 39.77510 & -122.67551 & 2002-04-28 & 3000-01-01 \\
HOPS & BK & BHE,BHN,BHZ & -- & NCEDC & 38.99349 & -123.07234 & 1994-10-21 & 2010-06-16 \\
HOPS & BK & BHE,BHN,BHZ & 0 & NCEDC & 38.99349 & -123.07234 & 2010-06-16 & 3000-01-01 \\
JCC & BK & BHE,BHN,BHZ & -- & NCEDC & 40.81745 & -124.02955 & 2001-04-11 & 2010-08-19 \\
JCC & BK & BHE,BHN,BHZ & 0 & NCEDC & 40.81745 & -124.02955 & 2010-08-19 & 3000-01-01 \\
KBN & NC & SHZ & -- & NCEDC & 39.89237 & -123.19503 & 1994-11-28 & 2011-10-27 \\
KBS & NC & SHZ & -- & NCEDC & 39.91719 & -123.59561 & 2002-10-17 & 2011-10-27 \\
KCPB & NC & HHE,HHN,HHZ & -- & NCEDC & 39.68631 & -123.58242 & 2002-10-17 & 3000-01-01 \\\
KCS & NC & SHZ & -- & NCEDC & 40.53791 & -123.51394 & 1994-11-28 & 2011-11-01 \\
KFP & NC & SHZ & -- & NCEDC & 39.63889 & -123.42514 & 1994-11-28 & 2011-10-27 \\
KHBB & NC & HHE,HHN,HHZ & -- & NCEDC & 40.65990 & -123.21966 & 2003-09-11 & 3000-01-01 \\
KHMB & NC & HHE,HHN,HHZ & -- & NCEDC & 40.87475 & -123.73259 & 2002-06-13 & 3000-01-01 \\
KIP & NC & SHZ & -- & NCEDC & 39.80841 & -123.48130 & 1994-11-28 & 2011-10-27 \\
KKP & NC & SHZ & -- & NCEDC & 40.14579 & -123.46965 & 1994-11-28 & 2011-11-01 \\
KOM & NC & SHZ & -- & NCEDC & 41.27872 & -123.45315 & 1994-11-28 & 2011-11-02 \\
KPP & NC & SHZ & -- & NCEDC & 40.34579 & -123.36328 & 1994-11-28 & 2011-11-01 \\
KRK & NC & SHZ & -- & NCEDC & 39.56315 & -123.18367 & 1994-11-28 & 2011-10-27 \\
KRMB & NC & HHE,HHN,HHZ & -- & NCEDC & 41.52296 & -123.90797 & 2001-06-16 & 3000-01-01 \\
KRP & NC & HHE,HHN,HHZ & -- & NCEDC & 41.15765 & -124.02330 & 2002-06-12 & 3000-01-01 \\
KSXB & NC & HHE,HHN,HHZ & -- & NCEDC & 41.83038 & -123.87688 & 2001-07-13 & 3000-01-01 \\
KTR & NC & SHZ & -- & NCEDC & 41.90847 & -123.37755 & 1994-11-28 & 2011-11-02 \\
LAM & NC & SHZ & -- & NCEDC & 41.60987 & -122.62559 & 1994-11-28 & 2011-09-30 \\
LBF & NC & SHZ & -- & NCEDC & 41.34707 & -121.89098 & 1994-11-28 & 2011-11-03 \\
LBK & NC & SHZ & -- & NCEDC & 41.08382 & -122.66731 & 1994-11-28 & 2011-12-08 \\
LBP & NC & SHZ & -- & NCEDC & 40.31671 & -122.88193 & 1994-11-28 & 2011-12-08 \\
LDB & NC & SHZ & -- & NCEDC & 40.43105 & -121.78632 & 1994-11-28 & 2011-12-08 \\
LGB & NC & SHZ & -- & NCEDC & 41.33418 & -122.18771 & 2001-10-20 & 2011-11-03 \\
LGP & NC & SHZ & -- & NCEDC & 40.91228 & -122.82949 & 1994-11-28 & 2011-12-08 \\
LPG & NC & SHZ & -- & NCEDC & 40.14514 & -122.68788 & 1994-11-28 & 2011-12-08 \\
LRB & NC & SHZ & -- & NCEDC & 40.14323 & -122.55772 & 1994-11-28 & 2011-12-08 \\
LRR & NC & SHZ & -- & NCEDC & 40.46630 & -121.62230 & 1994-11-28 & 2011-12-08 \\
LSF & NC & SHZ & -- & NCEDC & 40.65817 & -122.52371 & 1994-11-28 & 2011-12-08 \\
LSH & NC & SHZ & -- & NCEDC & 40.79294 & -122.03943 & 2002-03-28 & 2011-12-08 \\
LSR & NC & SHZ & -- & NCEDC & 41.10696 & -122.27062 & 1994-11-28 & 2011-12-08 \\
LTC & NC & SHZ & -- & NCEDC & 40.20842 & -122.12548 & 2002-03-28 & 2011-12-08 \\
LVR & NC & SHZ & -- & NCEDC & 40.03937 & -122.67250 & 1994-11-28 & 2011-12-08 \\
LWH & NC & SHZ & -- & NCEDC & 40.64180 & -121.94857 & 1994-11-28 & 2011-12-08 \\
WDC & BK & BHE,BHN,BHZ & -- & NCEDC & 40.57988 & -122.54113 & 1992-09-17 & 2011-05-06 \\
WDC & BK & BHE,BHN,BHZ & 0 & NCEDC & 40.57988 & -122.54113 & 2011-05-06 & 3000-01-01 \\
YBH & BK & BHE,BHN,BHZ & -- & NCEDC & 41.73204 & -122.71039 & 1993-07-24 & 2011-06-03 \\
YBH & BK & BHE,BHN,BHZ & 0 & NCEDC & 41.73204 & -122.71039 & 2011-06-03 & 3000-01-01 \\
\hline
\end{tabular}
\end{table}

\end{document}